### 2018-11-12 Sabanci_talk/Sabanci_talk.tex
New figs, v2 fnished

@@

\title[Forecasting GRBs w/ GWs] % (optional, use only with long paper titles)
{Forecasting Gamma-ray Bursts with Gravitational Waves}

\subtitle{\todo{Add background image}} % (optional)

\author[Sarp Akcay]{Sarp Akcay\inst{1}\inst{2} }
\institute[FSU Jena - UCD] % (optional, but mostly needed)
{
  \inst{1}%
  FSU Jena %and University College Dublin
  \inst{2}%
  University College Dublin
  }

%\vspace{-3cm}
%\titlegraphic{
%  \vspace{1.75cm}  \includegraphics[width=1.0cm]{figs/ucd.png}%\hspace*{1cm}~\includegraphics[height=1cm]{figs/missing_inch.jpg} \hspace*{1cm}  \includegraphics[width=1.0cm]{figs/ucd.png}
%}
\date[Sabanci University]{14 November 2018}

@@

\newcommand{\Mag}{\textcolor{magenta}}
\newcommand{\Red}{\textcolor{red}}
\newcommand{\Blue}{\textcolor{blue}}
\renewcommand{\c}{\cos}
\renewcommand{\t}{\theta}
\renewcommand{\b}{\bar}
\newcommand{\f}{\frac}
\newcommand{\bt}{\beta}
%\newcommand{\s}{\sin}
\newcommand{\ph}{\phi}
\newcommand{\g}{\gamma}
\newcommand{\nn}{\nonumber}
\newcommand{\la}{\lambda}
\newcommand{\al}{\alpha}
\newcommand{\La}{\Lambda}
\newcommand{\el}{\ell}
\newcommand{\h}{\hat}
\newcommand{\mrm}{\mathrm}
\newcommand{\ord}{\mathcal{O}}
\newcommand{\F}{\mathcal{F}}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\ba}{\begin{eqnarray}}
\newcommand{\ea}{\end{eqnarray}}
\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\bef}{\begin{frame}}
\newcommand{\ef}{\end{frame}}
%\newcommand{\h}{\bar{h}}
\newcommand{\bs}{\begin{small}}
\newcommand{\es}{\end{small}}
\newcommand{\parallelsum}{\mathbin{\!/\mkern-5mu/\!}}
\newcommand{\Lim}[1]{\raisebox{0.5ex}{\scalebox{0.8}{$\displaystyle \lim_{#1}\;$}}}
\newcommand\T{\rule{0pt}{2.6ex}}       % Top strut for vertical spacing in tables
\newcommand\B{\rule[-1.2ex]{0pt}{0pt}} % Bottom strut
\newcommand*\circled[1]{\tikz[baseline=(char.base)]{\node[shape=circle,draw,inner sep=2pt] (char) {#1};}}
%------------------ for coloured boxes in math modes --------------------
% Syntax: \colorboxed[<color model>]{<color specification>}{<math formula>}
\newcommand*{\colorboxed}{}
\def\colorboxed#1#{%
  \colorboxedAux{#1}%
}
\newcommand*{\colorboxedAux}[3]{%
  % #1: optional argument for color model
  % #2: color specification
  % #3: formula
  \begingroup
    \colorlet{cb@saved}{.}%
    \color#1{#2}%
    \boxed{%
      \color{cb@saved}%
      #3%
    }%
  \endgroup
}
%%-------------------------------------------------------------------------

@@

\begin{document}
%{
%\usebackgroundtemplate{\includegraphics[width=1\paperwidth,height=\paperheight]{../Figures/ET_strains_redshifted_v2.pdf}}
\begin{frame}
 \titlepage
\end{frame}
%}

\begin{frame}{Outline}
\begin{itemize}
 \item Motivation
 \item Gamma-ray bursts
 \item Gravitational waves
 \item Binary neutron star inspirals
 \item Gravitational-wave detectors
 \item Future detectors, event rates, \red{early warning} prospects
\end{itemize}
\todo{finish this slide remove later if short on time}

@@

\begin{frame}{Motivation: Multi-messenger Astronomy}
  \begin{itemize}
    \item \emph{Multi}: two fundamentally different waves from common events.
    \item[]\quad Gravitational waves (GWs): ``hearing''  the universe.
    \item[]\quad Electromagnetic waves (EWs): ``seeing'' the universe.
    \item \red{GW170817}-\bl{GRB170817A}-\textcolor{brown}{AT2017gfo}: the ``event'' \\
    \quad inspiral + merger of two neutron stars (GWs) \\
    \ \ = short-hard gamma-ray burst + kilonova (EWs)
    \item[] Broad-band in gravitational waves: 10Hz - 2000Hz
    \item[] Ultra broad-band in electromagnetic waves: gamma rays to radio.
    \item A simple question \\
    {\small ``What can we offer to the astronomy community with future GW detections?''}
    \item[] \red{GRB Forecasts!}
  \end{itemize}
\end{frame}

\begin{frame}{Gamma-ray bursts}{Most energetic explosions in the universe}
\begin{small}
    \begin{itemize}
      \item Discovered by Vela satellites in 1967
      \onslide<2->\item Non-thermal spectrum: $10-10^4\,$keV
      \item Highly energetic $10^{48}-10^{52}\,$ergs 
      \onslide<3->{\item 1000s detected, $\sim 1$ per day}
      \onslide<4->\item Far away: $z\approx 9.4 \sim 13\,$billion light years!
      \onslide<5->\item Afterglow from X-rays to radio \\
      over timescales of hours to months
      \onslide<6->\item Two types: short and long
      \onslide<7->{\item Short GRBs:\\ 
      mergers of binary neutron stars\\
      \red{GW170817}-\bl{GRB170817A}}
    \end{itemize}
  \end{small}
  \begin{tikzpicture}[overlay,remember picture]
      \onslide<2>\node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(3.9cm,1.4cm)}]current page.center) {\includegraphics[height=3.5cm]{figs/GRB_spectrum.jpg}};    
      \node[anchor=center,scale=0.5,opacity=1] at ([shift={(4.5cm,1.cm)}]current page.center) { Mon.Not.Roy.Astron.Soc. 433 (2013) 2739};
      \onslide<3>\node (img1 )[anchor=center,scale=0.8,opacity=1] at ([shift={(3.8cm,0.9cm)}]current page.center) {\includegraphics[height=3.5cm]{figs/Fermi_GRB_sky2.png}};  
      \node[anchor=center,scale=0.5,opacity=1] at ([shift={(5.5cm,1.9cm)}]current page.center) {NASA/Fermi};
      \onslide<4>\node (img2 )[anchor=center,scale=1,opacity=1] at ([shift={(3.6cm,-0.15cm)}]current page.center) {\includegraphics[height=3.8cm]{figs/GRB_vs_redshift.png}};  % from:1401.5069
      \node[anchor=center,scale=0.5,opacity=1] at ([shift={(4.cm,1.4cm)}]current page.center) {Astrophys.J.Suppl. 211 (2014) 12};
      \onslide<5>\node (img3 )[anchor=center,scale=1,opacity=1] at ([shift={(3.4cm,-1.5cm)}]current page.center) {\includegraphics[height=3.5cm]{figs/GRB_afterglow.jpg}}; 
      \onslide<6->\node (img4 )[anchor=center,scale=1,opacity=1] at ([shift={(3.2cm,-2.5cm)}]current page.center) {\includegraphics[height=3.5cm]{figs/GRB_burst_durations.jpg}};
      \node[anchor=center,scale=0.5,opacity=1] at ([shift={(2.cm,-1.2cm)}]current page.center) {NASA/GSFC};
    \end{tikzpicture}
\end{frame}

\begin{frame}{Neutron Stars}{Most extreme stars in the universe}  
  \begin{small}
    \begin{itemize}
      \item Core-collapse supernova remnants
      \item Mass: $\sim M_\odot$ to $3 M_\odot$
      \item Very \red{compact}: radius $\sim 9$ to 15\,km
      \onslide<2->\item High spins: up to 700 times/second
      \onslide<3->\item Extremely dense: teaspoon $\sim 10^9\,$tons
      \onslide<3->{\item Large magnetic fields: $10^4$ to $10^{11}\,$T}
      \item Hot: surface temperature $10^6\,$K
      \onslide<4->\item Equation of state \\
      Piecewise polytropes
      \[ p = K_i \rho^{\Gamma_i},\quad (\rho_{i-1}\le \rho \le \rho_i) \hspace{8cm}\]
    \end{itemize}
  \end{small}
  \begin{tikzpicture}[overlay,remember picture]
      \onslide<1>
	\node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(4.2cm,1.5cm)}]current page.center) {\includegraphics[height=3.2cm]{figs/Core_collapseSXS.jpg}};    
	\node[anchor=center,scale=1,opacity=1] at ([shift={(5.4cm,2.9cm)}]current page.center) {\textcolor{white}{SXS}};
      \onslide<2>\node (img1 )[anchor=center,scale=0.8,opacity=1] at ([shift={(3.8cm,0.9cm)}]current page.center) {\includegraphics[height=6cm]{figs/Millisecond_NS.png}}; 
      \node[anchor=center,scale=0.5,opacity=1] at ([shift={(4.8cm,-1.cm)}]current page.center) {Class.Quant.Grav. 30 (2013) 224003};
      \onslide<3>\node (img2 )[anchor=center,scale=0.8,opacity=1] at ([shift={(3.6cm,0.0cm)}]current page.center) {\includegraphics[height=4cm]{figs/NS_cross_section.png}}; 
      % Use Fig 3 of http://science.sciencemag.org/content/304/5670/536/tab-figures-data ?
      \node[anchor=center,scale=0.6,opacity=1] at ([shift={(4.6cm,-1.5cm)}]current page.center) {Robert Schulze};
       \onslide<4>\node (img3 )[anchor=center,scale=0.8,opacity=1] at ([shift={(2.7cm,-2.5cm)}]current page.center) {\includegraphics[height=5cm]{figs/NS_EOS_Lattimer2012.png}}; 
      \node[anchor=center,scale=0.6,opacity=1] at ([shift={(4.cm,-3.8cm)}]current page.center) {Lattimer arXiv:1305.3510};
    \end{tikzpicture}
\end{frame}

@@

 \begin{frame}{Making gravitational waves}{non-spherical, accelerated motion}
  \alert{Inspirals}, GRBs, bumps on neutron stars, supernovae, the Big Bang. \\%  (BICEP 2).\\
  Gravitational-wave luminosity: \alert{quadrupole} radiation (leading-order)
  \begin{center} $ \boxed{\dot{E}=L_\text{GW} = \f{G}{c^5} \langle (\partial_t^3 \bar{Q}_{ij})^2 \rangle}$\\
  \vspace{2mm}
  \begin{small}$ \quad \bar{Q}_{ij} = Q_{ij}-\frac{1}{3}\delta_{ij}Q^k_k, \quad Q^{ij} = \int \rho(t,{\mathbf x})x^i x^j d^3 x $ \end{small} \end{center}
  \textcolor{white}{{\bf Binary systems:} 2 points masses in circular orbit.}
  \begin{tikzpicture}[overlay,remember picture]
    \node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(-3.8cm,-2.7cm)}]current page.center) {\includegraphics[height=2.9cm]{figs/white_box.png}};
  \end{tikzpicture}
\end{frame}

\begin{frame}{Making gravitational waves}{non-spherical, accelerated motion}
  \alert{Inspirals}, GRBs, bumps on neutron stars, supernovae, the Big Bang. \\%  (BICEP 2).\\
  Gravitational-wave luminosity: \alert{quadrupole} radiation (leading-order)
  \begin{center} $ \boxed{\dot{E}=L_\text{GW} = \f{G}{c^5} \langle (\partial_t^3 \bar{Q}_{ij})^2 \rangle}$\\
  \vspace{2mm}
  \begin{small}$ \quad \bar{Q}_{ij} = Q_{ij}-\frac{1}{3}\delta_{ij}Q^k_k, \quad Q^{ij} = \int \rho(t,{\mathbf x})x^i x^j d^3 x $ \end{small} \end{center}
  {\bf Binary systems:} 2 points masses in circular orbit.
  \begin{tikzpicture}[overlay,remember picture]
    \node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(-3.8cm,-2.7cm)}]current page.center) {\animategraphics[loop,controls,height=2.9cm]{12}{Circ_Orb-}{0}{20}};
    %\node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(-3.8cm,-2.7cm)}]current page.center) {\includegraphics[height=2.9cm]{figs/white_box.png}};
    \onslide<2->\node (t1 )[anchor=center,scale=1,opacity=1] at ([shift={(2cm,-1.3cm)}]current page.center) {$\mu=m_1 m_2/M^2,\ x(t) = r \cos\Omega t, \ y(t) = r \sin\Omega t $};
    \onslide<3->\node (t2 )[anchor=center,scale=1,opacity=1] at ([shift={(2.25cm,-2.2cm)}]current page.center) {$ Q^{ij} =  2 M\, x^i x^j, \qquad \boxed{\dot{E} = \f{32}{5}\f{G\mu^2}{c^5}r^4 \Omega^6}$};
    \onslide<4->\node (t3 )[anchor=center,scale=1,opacity=1] at ([shift={(2.5cm,-3.2cm)}]current page.center) {Enhancement due to \bl{$e$} (Peters \& Mathews).};
    \node (t3 )[anchor=center,scale=1,opacity=1] at ([shift={(2cm,-4cm)}]current page.center) {$ \bl{\f{\left(1 + \f{73}{24}e^2+ \f{37}{96} e^4 \right)}{(1-e^2)^{7/2}}}$};
  \end{tikzpicture}
\end{frame}

@@

\begin{frame}{Compact Binary Inspirals}
  \vspace{-3mm}
  Notation:
  \begin{itemize}
    \item $\red{f} $: GW frequency, \hspace{1.5cm} $\Longrightarrow\quad \omega\equiv2\pi f = 2\Omega$.
    %\item $\omega\equiv2\pi f = 2\Omega$: GW angular frequency.
    \item $\red{t}$: observer/detector time \quad$ \Longrightarrow\quad\dot{Y}= dY/dt$.
  \end{itemize}
  \onslide<2->
  {
    Assumptions:
    \begin{itemize}
      \item \bl{Quasi-circular} orbits: $e \ll 1$ and $\tfrac{|\dot{r}|}{r\Omega} <10^{-3} $ at $f=100\,$Hz.
      \item \bl{Quadrupolar} GWs only: higher modes suppressed by $\tfrac{m_1-m_2}{M}\tfrac{v^2}{c^2}$
      \item Use Newtonian/post-Newtonian theory ($\le$3.5{\bf PN}).
      \item \red{Terminate} inspiral at Schwarzschild \red{ISCO} 
      \begin{footnotesize}
      \begin{align*}
      f_\text{ISCO} &= \f{c^3}{6^{3/2}\pi G M} \simeq \bl{1571}\,\text{Hz}\, \left(\f{2.8M_\odot}{M}\right)\hspace{2cm}\\ 
      r_\text{ISCO} &\simeq 25\,\text{km}\left(\frac{M}{2.8 M_\odot}\right)
      \end{align*}
      \end{footnotesize}
    \end{itemize}
   } 
  %
  \onslide<3->
  {
    \vspace{1mm}
    For binary neutron stars ({\bf BNS})
    \begin{itemize}
      \item Point masses: tidal effects for $f \gtrsim 100\,$Hz
      \item Set $m_1 =m_2 = 1.4 M_\odot$, constraints: $1.2 M_\odot \lesssim m_\text{NS}\lesssim 1.6M_\odot$.
      \item Neglect spins: spin-orbit $\tfrac{\dot{E}_\text{SO}}{\dot{E}} \sim \tfrac{v_s}{c}\tfrac{v}{c}\tfrac{R_\text{NS}}{r}$, spin-spin $\sim \tfrac{v^4}{c^4}$
    \end{itemize}
  }
  \onslide<2->
  {
    \begin{tikzpicture}[overlay,remember picture]
    \node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(4.4cm,-1.5cm)}]current page.center) {\includegraphics[height=3.5cm]{figs/OrbitPotential.pdf}};    
    \end{tikzpicture}
  }  
\end{frame}

>>>>

%\setlength{\droptitle}{10em}

@@

\title[Forecasting GRBs with GWs] % (optional, use only with long paper titles)
{\vspace{0mm}Forecasting Gamma-ray Bursts with Gravitational Waves}

\subtitle{} % (optional)
%\vspace{2cm}
\author[Sarp Ak\c{c}ay]{\vspace{-2cm}Sarp Ak\c cay\inst{1}\inst{2} }
\institute[FSU Jena - UCD] % (optional, but mostly needed)
{\vspace{-2.25cm}
  \inst{1}%
  FSU Jena %and University College Dublin
  \inst{2}%
  University College Dublin
  }

\date[Sabanc\i~Universitesi]{\vspace{-2.5cm}14 November 2018}
%\vspace{-3cm}
%\titlegraphic{
%  \vspace{1.75cm}  \includegraphics[width=1.0cm]{figs/ucd.png}%\hspace*{1cm}~\includegraphics[height=1cm]{figs/missing_inch.jpg} \hspace*{1cm}  \includegraphics[width=1.0cm]{figs/ucd.png}
%}

@@

\newcommand{\Mag}{\textcolor{magenta}}
\newcommand{\Red}{\textcolor{red}}
\newcommand{\Blue}{\textcolor{blue}}
%\renewcommand{\c}{\cos}  \c{c} gives cecedille
\renewcommand{\t}{\theta}
\renewcommand{\b}{\bar}
\newcommand{\f}{\frac}
\newcommand{\bt}{\beta}
%\newcommand{\s}{\sin}
\newcommand{\ph}{\phi}
\newcommand{\g}{\gamma}
\newcommand{\nn}{\nonumber}
\newcommand{\la}{\lambda}
\newcommand{\al}{\alpha}
\newcommand{\La}{\Lambda}
\newcommand{\el}{\ell}
\newcommand{\h}{\hat}
\newcommand{\mrm}{\mathrm}
\newcommand{\ord}{\mathcal{O}}
\newcommand{\F}{\mathcal{F}}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\ba}{\begin{eqnarray}}
\newcommand{\ea}{\end{eqnarray}}
\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\bef}{\begin{frame}}
\newcommand{\ef}{\end{frame}}
%\newcommand{\h}{\bar{h}}
\newcommand{\bs}{\begin{small}}
\newcommand{\es}{\end{small}}
\newcommand{\parallelsum}{\mathbin{\!/\mkern-5mu/\!}}
\newcommand{\Lim}[1]{\raisebox{0.5ex}{\scalebox{0.8}{$\displaystyle \lim_{#1}\;$}}}
\newcommand\T{\rule{0pt}{2.6ex}}       % Top strut for vertical spacing in tables
\newcommand\B{\rule[-1.2ex]{0pt}{0pt}} % Bottom strut
\newcommand*\circled[1]{\tikz[baseline=(char.base)]{\node[shape=circle,draw,inner sep=2pt] (char) {#1};}}
%------------------ for coloured boxes in math modes --------------------
% Syntax: \colorboxed[<color model>]{<color specification>}{<math formula>}
\newcommand*{\colorboxed}{}
\def\colorboxed#1#{%
  \colorboxedAux{#1}%
}
\newcommand*{\colorboxedAux}[3]{%
  % #1: optional argument for color model
  % #2: color specification
  % #3: formula
  \begingroup
    \colorlet{cb@saved}{.}%
    \color#1{#2}%
    \boxed{%
      \color{cb@saved}%
      #3%
    }%
  \endgroup
}
%%-------------------------------------------------------------------------

@@

\begin{document}
{
  \usebackgroundtemplate{\includegraphics[width=1\paperwidth,height=\paperheight]{figs/kilonova.jpg}}
  % image credit [University of Warwick/Mark Garlick]
  \begin{frame}
  \vspace{-2cm}
    \titlepage
  \end{frame}
} % these brackets are needed to insure that the background image is used ONLY in the title slide

\begin{frame}{Outline}
\begin{itemize}
 \item Motivation 
 \item \bl{Gamma-ray bursts}
 \item Neutron stars
 \item \bl{Gravitational waves}
 \item Compact binary \red{inspirals}
 \item Gravitational-wave detectors
 \item Gravitational-wave astronomy in the 2020s and the 2030s
 \item \red{Forecasting} prospects, event rates etc.
\end{itemize}

@@

\begin{frame}{Motivation: Multi-messenger Astronomy}
  \begin{itemize}
    \onslide<1->
    \item \red{\emph{Multi}}: two fundamentally different waves from common events.
    \item[]\quad Gravitational waves ({\bf GW}s): ``hearing''  the universe.
    \item[]\quad Electromagnetic waves (EWs): ``seeing'' the universe.
    \onslide<2->\item \red{GW170817}-\bl{GRB170817A}-\textcolor{brown}{AT2017gfo}: the ``event'' \\
    \quad inspiral + merger of two neutron stars (GWs) \\
    \ \ \ = short-hard gamma-ray burst + kilonova (EWs)
    \item[] Broad-band in gravitational waves: 20Hz - 2000Hz
    \item[] Ultra broad-band in electromagnetic waves: gamma rays to radio.
    \onslide<3->\item A simple question: \\
    {\small ``What can we offer to the astronomy community with future GW detections?''}
    \item[] \red{GRB/kilonova Forecasts!}
    \item[] Observe NOT 11 hours \underline{after} the merger, but ``live''
  \end{itemize}
\end{frame}

\begin{frame}{Gamma-ray bursts}{Most energetic explosions in the universe}
\begin{small}
    \begin{itemize}
      \item Discovered by Vela satellites in \red{1967}
      \onslide<2->\item Non-thermal spectrum: $10-10^4\,$\red{keV}
      \item Highly energetic $10^{48}-10^{\red{52}}\,$ergs 
      \onslide<3->{\item \bl{1000s} detected, $\sim 1$ per day}
      \onslide<4->\item Far away: $z\approx 1 \sim 8\,$billion years! %$9.4\sim 13\,$billion light years!
      \onslide<5->\item Afterglow from X-rays to radio \\
      over timescales of hours to months
      \onslide<6->\item \bl{Two types}: \red{short} and long
      \onslide<7->{\item Short GRBs:\\ 
      mergers of binary neutron stars (1989)\\
      \red{GW170817}-\bl{GRB170817A}}
    \end{itemize}
  \end{small}
  \begin{tikzpicture}[overlay,remember picture]
      \onslide<2>\node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(3.9cm,1.4cm)}]current page.center) {\includegraphics[height=3.5cm]{figs/GRB_spectrum.jpg}};    
      \node[anchor=center,scale=0.5,opacity=1] at ([shift={(4.5cm,1.cm)}]current page.center) { Mon.Not.Roy.Astron.Soc. 433 (2013) 2739};
      \onslide<3>\node (img1 )[anchor=center,scale=0.8,opacity=1] at ([shift={(3.8cm,0.9cm)}]current page.center) {\includegraphics[height=3.5cm]{figs/Fermi_GRB_sky2.png}};  
      \node[anchor=center,scale=0.5,opacity=1] at ([shift={(5.5cm,1.9cm)}]current page.center) {NASA/Fermi};
      \onslide<4>\node (img2 )[anchor=center,scale=1,opacity=1] at ([shift={(3.6cm,-0.15cm)}]current page.center) {\includegraphics[height=3.8cm]{figs/GRB_vs_redshift.png}};  % from:1401.5069
      \node[anchor=center,scale=0.5,opacity=1] at ([shift={(4.cm,1.4cm)}]current page.center) {Astrophys.J.Suppl. 211 (2014) 12};
      \onslide<5>\node (img3 )[anchor=center,scale=1,opacity=1] at ([shift={(3.4cm,-1.5cm)}]current page.center) {\includegraphics[height=3.5cm]{figs/GRB_afterglow.jpg}}; 
      \onslide<6->\node (img4 )[anchor=center,scale=1,opacity=1] at ([shift={(3.2cm,-2.5cm)}]current page.center) {\includegraphics[height=3.5cm]{figs/GRB_burst_durations.jpg}};
      \node[anchor=center,scale=0.5,opacity=1] at ([shift={(2.cm,-1.2cm)}]current page.center) {NASA/GSFC};
    \end{tikzpicture}
\end{frame}

\begin{frame}{Neutron Stars}{Most extreme stars in the universe}  
  \begin{small}
    \begin{itemize}
      \item Core-collapse supernova remnants
      \item Mass: $\sim $ \red{1} to \red{3} solar masses ($M_\odot$)
      \item Very \red{compact}: radius $\sim 9$ to 15\,km
      \onslide<2->\item High spins: up to 700 times/second
      \onslide<3->\item Extremely \bl{dense}: teaspoon $\sim 10^9\,$tons
      \onslide<3->{\item Large magnetic fields: $10^4$ to $10^{\red{11}}\,$T}
      \onslide<4->\item Hot: surface temperature $10^5 - 10^6\,$K
      \onslide<5->\item They can come in \red{pairs} (binary) \\
      \onslide<6->Orbit \red{decays} over time due\\
      $\Longrightarrow$ \underline{Gravitational wave emission}!\\
      Observations \red{match} theoretical \\predictions very well
      %Equation of state \\
      %Piecewise polytropes
      %\[ p = K_i \rho^{\Gamma_i},\quad (\rho_{i-1}\le \rho \le \rho_i) \hspace{8cm}\]
    \end{itemize}
  \end{small}
  \begin{tikzpicture}[overlay,remember picture]
      \onslide<1>
	\node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(4.2cm,1.5cm)}]current page.center) {\includegraphics[height=3.2cm]{figs/Core_collapseSXS.jpg}};    
	\node[anchor=center,scale=1,opacity=1] at ([shift={(5.4cm,2.9cm)}]current page.center) {\textcolor{white}{SXS}};
      \onslide<2>\node (img1 )[anchor=center,scale=0.8,opacity=1] at ([shift={(3.8cm,0.9cm)}]current page.center) {\includegraphics[height=6cm]{figs/Millisecond_NS2.png}}; 
      \node[anchor=center,scale=0.5,opacity=1] at ([shift={(4.8cm,3.2cm)}]current page.center) {Class.Quant.Grav. 30 (2013) 224003};
      \onslide<3>\node (img2 )[anchor=center,scale=0.8,opacity=1] at ([shift={(3.6cm,0.0cm)}]current page.center) {\includegraphics[height=4cm]{figs/NS_cross_section.png}}; 
      % Use Fig 3 of http://science.sciencemag.org/content/304/5670/536/tab-figures-data ?
      \node[anchor=center,scale=0.6,opacity=1] at ([shift={(4.6cm,-1.5cm)}]current page.center) {Robert Schulze};
      \onslide<4>\node (img2 )[anchor=center,scale=0.8,opacity=1] at ([shift={(3.6cm,0.0cm)}]current page.center) {\includegraphics[height=5cm]{figs/NS_temp.jpg}}; 
      %
      \node[anchor=center,scale=0.6,opacity=1] at ([shift={(4.6cm,2.2cm)}]current page.center) {Science 304, 5670, pp. 536-542};
       \onslide<5->\node (img3 )[anchor=center,scale=0.8,opacity=1] at ([shift={(3.cm,-2.8cm)}]current page.center){\includegraphics[height=3.8cm]{figs/HT_pulsar.jpg}};
       \onslide<6>\node (img3 )[anchor=center,scale=0.8,opacity=1] at ([shift={(3.2cm,0.9cm)}]current page.center){\includegraphics[height=5cm]{figs/HT_decay.jpg}}; 
       \node[anchor=center,scale=0.5,opacity=0.7] at ([shift={(2.5cm,-0.5cm)}]current page.center) {ApJ 829: 55 (2016)};
       %{\includegraphics[height=5cm]{figs/NS_EOS_Lattimer2012.png}}; 
      %\node[anchor=center,scale=0.6,opacity=1] at ([shift={(4.cm,-3.8cm)}]current page.center) {Lattimer arXiv:1305.3510};
    \end{tikzpicture}
\end{frame}

@@

 \begin{frame}{Making gravitational waves}{non-spherical, accelerated motion}
  \alert{Inspirals}, GRBs, bumps on neutron stars, supernovae, the Big Bang. \\%  (BICEP 2).\\
  Gravitational-wave emission: \alert{quadrupole} radiation (leading-order)
  \begin{center} $ \boxed{\dot{E}=L_\text{GW} = \f{G}{c^5} \langle (\partial_t^3 \bar{Q}_{ij})^2 \rangle}$\\
  \vspace{2mm}
  \begin{small}$ \quad \bar{Q}_{ij} = Q_{ij}-\frac{1}{3}\delta_{ij}Q^k_k, \quad Q^{ij} = \int \rho(t,{\mathbf x})x^i x^j d^3 x $ \end{small} \end{center}
  \textcolor{white}{{\bf Binary systems:} 2 points masses in circular orbit.}
  \begin{tikzpicture}[overlay,remember picture]
    \node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(-3.8cm,-2.7cm)}]current page.center) {\includegraphics[height=2.9cm]{figs/white_box.png}};
  \end{tikzpicture}
\end{frame}

\begin{frame}{Making gravitational waves}{non-spherical, accelerated motion}
  \alert{Inspirals}, GRBs, bumps on neutron stars, supernovae, the Big Bang. \\%  (BICEP 2).\\
  Gravitational-wave emission: \alert{quadrupole} radiation (leading-order)
  \begin{center} $ \boxed{\dot{E}=L_\text{GW} = \f{G}{c^5} \langle (\partial_t^3 \bar{Q}_{ij})^2 \rangle}$\\
  \vspace{2mm}
  \begin{small}$ \quad \bar{Q}_{ij} = Q_{ij}-\frac{1}{3}\delta_{ij}Q^k_k, \quad Q^{ij} = \int \rho(t,{\mathbf x})x^i x^j d^3 x $ \end{small} \end{center}
  {\bf Binary systems:} 2 points masses in circular orbit.
  \begin{tikzpicture}[overlay,remember picture]
    \node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(-3.8cm,-2.7cm)}]current page.center) {\animategraphics[loop,controls,height=2.9cm]{12}{Circ_Orb-}{0}{20}};
    %\node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(-3.8cm,-2.7cm)}]current page.center) {\includegraphics[height=2.9cm]{figs/white_box.png}};
    \onslide<2->\node (t1 )[anchor=center,scale=1,opacity=1] at ([shift={(2cm,-1.3cm)}]current page.center) {$\mu=m_1 m_2/M^2,\ x(t) = r \cos\Omega t, \ y(t) = r \sin\Omega t $};
    \onslide<3->\node (t2 )[anchor=center,scale=1,opacity=1] at ([shift={(2.25cm,-2.2cm)}]current page.center) {$ Q^{ij} =  \mu\, x^i x^j, \qquad \boxed{\dot{E} = \f{32}{5}\f{G\mu^2}{c^5}r^4 \Omega^6}$};
    \onslide<4->\node (t3 )[anchor=center,scale=1,opacity=1] at ([shift={(2.5cm,-3.2cm)}]current page.center) {Enhancement due to eccentricity \bl{$e$}};
    \node (t3 )[anchor=center,scale=1,opacity=1] at ([shift={(2cm,-4cm)}]current page.center) {$ \bl{\f{\left(1 + \f{73}{24}e^2+ \f{37}{96} e^4 \right)}{(1-e^2)^{7/2}}}$};
  \end{tikzpicture}
\end{frame}

@@

\begin{frame}{Compact Binary Inspirals}
  \vspace{-4mm}
  Notation:
  \begin{itemize}
    \item $\red{f} $: GW frequency, \hspace{1.5cm} $\Longrightarrow\quad \omega\equiv2\pi f = 2\Omega$.
    %\item $\omega\equiv2\pi f = 2\Omega$: GW angular frequency.
    \item $\red{t}$: observer/detector time \quad$ \Longrightarrow\quad\dot{Y}= dY/dt$.
  \end{itemize}
  \onslide<2->
  {
    Assumptions:
    \begin{itemize}
      \item \bl{Quasi-circular} orbits: $e \ll 1$ and $\tfrac{|\dot{r}|}{r\Omega} <10^{-3} $ %at $f=100\,$Hz.
      \item \bl{Quadrupolar} GWs only: higher modes suppressed by $\tfrac{m_1-m_2}{M}\tfrac{v^2}{c^2}$
      \item Use Newtonian/post-Newtonian theory ($\le$3.5{\bf PN}).
      \vspace{0.5mm}
      \item \red{Terminate} inspiral at Schwarzschild \red{ISCO} 
      \begin{footnotesize}
      \begin{align*}
      f_\text{ISCO} &= \f{c^3}{\red{6}^{3/2}\pi G M} \simeq \bl{1571}\,\text{Hz}\, \left(\f{2.8M_\odot}{M}\right)\hspace{2cm}\\ 
      r_\text{ISCO} &\simeq 25\,\text{km}\left(\frac{M}{2.8 M_\odot}\right)
      \end{align*}
      \end{footnotesize}
    \end{itemize}
   } 
  %
  \onslide<3->
  {
    \vspace{1mm}
    For binary neutron stars ({\bf BNS})
    \begin{itemize}
      \item Point masses: tidal effects beyond $f \gtrsim 100\,$Hz
      \item Set $m_1 =m_2 = 1.4 M_\odot$, constraints: $1.2 M_\odot \lesssim m_\text{NS}\lesssim 1.6M_\odot$.
      \item Neglect spins: spin-orbit $\tfrac{\dot{E}_\text{SO}}{\dot{E}} \sim \tfrac{v_s}{c}\tfrac{v}{c}\tfrac{R_\text{NS}}{r}$, spin-spin $\sim \tfrac{v^4}{c^4}$
    \end{itemize}
  }
  \onslide<2->
  {
    \begin{tikzpicture}[overlay,remember picture]
    \node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(4.4cm,-1.5cm)}]current page.center) {\includegraphics[height=3.5cm]{figs/OrbitPotential.pdf}};    
    \end{tikzpicture}
  }  
\end{frame}
